---
  title: "paired WSR simulation : multivariate exponential and weibull"
author: "민태원, 권대영"
date: "`r Sys.Date()`"
output: html_document
---
  
#### load `lcmix` package which includes `rmvnorm` and `rmvexp`.
  
  
```{r include=F}
#install.packages("nnls") #install before download lcmix
#install.packages("lcmix", repos="http://R-Forge.R-project.org")
#install.packages("flexsurv")
#library(lcmix)
library(combinat)
library(flexsurv)
library(gtools)
source("mvfunctions.R")
```

## (1) Specification

* Number of simulations = 1000
* one-sided, alpha = 0.05
* $\Delta_x = \frac{\mu_{x2} - \mu_{x1}}{\sqrt{2}\sigma_{x1}} \in \{0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6\}$ is the primary outcome of interest ($\mu_{x1} = 0$, $\sigma_{x1} = 1$)
* $\begin{pmatrix} X_1 \\ X_2 \end{pmatrix} \sim N \begin{pmatrix} \begin{pmatrix} 0\\\sqrt{2} \Delta_x \end{pmatrix}, \begin{pmatrix} \sigma^2_{x1} & \rho_x\sigma_{x1}\sigma_{x2} \\ \rho_x\sigma_{x1}\sigma_{x2} & \sigma^2_{x2} \end{pmatrix} \end{pmatrix}$, $\rho_x=0.8$, $\sigma_{x1}=\sigma_{x2}=1$
  * $p_i = E(\delta_{ij}) = \mathrm{Pr}(\delta_{ij}=1) = \mathrm{Pr}(t_{ij} \leq T)$ is the probability of truncated by death at observed time T for group i
* $q_i = 1-p_i = \mathrm{Pr}(t_{ij} > T), \quad i = 1, 2$ is the probability of survival at observed time T
* $q_2 \in \{0.6, 0.8\}, \quad T = 3$
  * $\text{HR (Hazard Ratio)} \in \{1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0  \}$
  
  
  
  
## (2) Power computation by Simulation

### (a) death follows exponential distribution

* $\begin{pmatrix} t_1 \\ t_2 \end{pmatrix} \sim Exp\left( \begin{pmatrix} \lambda_{1}\\ \lambda_{2} \end{pmatrix}, \rho_t \right), \quad \rho_t = 0.8$ 
* marginal pdf $f(t; \lambda) = \lambda e^{-\lambda t} ,\quad t \ge 0$
* $q_i = exp(-\lambda_i T), \quad i=1,2, \quad T=3, \quad q_2 \in \{0.6, 0.8\}$
* $h_{i}(t) = \lambda_i$
* $\text{HR} = h_{1}(t)/h_{2}(t) = \lambda_1 / \lambda_2$
  
  
#### function that generate data from `rmvexp` and define parameters
  
```{r}
#Define parameters 
Delta = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6)
q2 = c(0.6, 0.8, 0.9)
HR = c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)
time = 3
N = 3000
n = 50
rho_X = 0.8
rho_t = 0.8
alpha = 0.05

simulate_exp <- function(n, q2, HR, Delta, time, rho_X, rho_t){
  lambda2 = -1 / time * log(q2)
  lambda1 = lambda2 * HR
  #q1 = exp(-lambda1 * t)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvexp(n, rate=c(lambda1, lambda2), corr=corr_t)
  t1 = t[,1]
  t2 = t[,2]
  
  corr_X = matrix(c(1, rho_X, rho_X, 1), nrow = 2)
  X = rmvnorm(n, mean = c(0, 2^0.5 * Delta), cov = corr_X)
  X1 = X[,1]
  X2 = X[,2]
  
  eta = min(X) - 1 - time - 2 *(max(X)-min(X)) # in order to be worst
  distmin = min(abs(X1-X2))/time/10 # less than minimum difference between X1 and X2 divided by time
  
  X1_tilde = X1 * (t1 >= time) + (eta + t1*distmin) * (t1 < time)
  X2_tilde = X2 * (t2 >= time) + (eta + t2*distmin) * (t2 < time)
  
  X1_survivorOnly = X1[t1 >= time & t2 >= time]
  X2_survivorOnly = X2[t1 >= time & t2 >= time]
  
  if (length(X1_survivorOnly) == 0) {
    return (c(-1,-1,-1))
  }
  return (c(wilcox.test(X1_tilde, X2_tilde, alternative="less", paired=T)$p.value,
            wilcox.test(X1_tilde, X2_tilde, alternative="less", paired=F)$p.value,
            wilcox.test(X1_survivorOnly, X2_survivorOnly, alternative="less", paired=T)$p.value))
}
```


#### Simulate(N = 3000, n = 50)

```{r}
power.exp = array(data = NA, dim = c(3, length(q2), length(Delta), length(HR)), 
                  dimnames = list(c("Paired Worst rank", "Non-paired Worst rank", "Survivors-only"),
                                  c("q2=60%", "q2=80%", "q2=90%"), 
                                  c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6), 
                                  c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)))
set.seed(123)
for (i in 1:length(q2)) {
  for (j in 1:length(Delta)) {
    print(Delta[j])
    for (k in 1:length(HR)) {
      count_wsr=0
      count_wmw=0
      count_survivor = 0
      total = 0
      for (l in c(1:N)){
        wsr_tests = simulate_exp(n, q2[i], HR[k], Delta[j], time, rho_X, rho_t)
        wsr_p = wsr_tests[1]
        wmw_p = wsr_tests[2]
        wsr_survivor_p = wsr_tests[3]
        if (wsr_p == -1) break;
        if (wsr_p <= 0.05){
          count_wsr = count_wsr + 1
        }
        if (wmw_p <= 0.05){
          count_wmw = count_wmw +1
        }
        if (wsr_survivor_p <= 0.05){
          count_survivor = count_survivor + 1
        }
        total = total + 1
      }
      power.exp[1, i, j, k] = round(count_wsr/total, 3)
      power.exp[2, i, j, k] = round(count_wmw/total, 3)
      power.exp[3, i, j, k] = round(count_survivor/total, 3)
    }
  }
}
```

#### Result

```{r}
wsr_q2low = power.exp[1,1,,]
wsr_q2high = power.exp[1,2,,]
#wsr_q2veryhigh = power.exp[1,3,,]

wmw_q2low = power.exp[2,1,,]
wmw_q2high = power.exp[2,2,,]
#wmw_q2veryhigh = power.exp[2,3,,]

survivoronly_q2low = power.exp[3,1,,]
survivoronly_q2high = power.exp[3,2,,]
#survivoronly_q2veryhigh = power.exp[3,3,,]

cat("Wilcoxon signed worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
wsr_q2low
cat("\nWilcoxon signed worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
wsr_q2high
#cat("\nWilcoxon signed worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.9\n")
#wsr_q2veryhigh
cat("\nWilcoxon-Mann-Whitney worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
wmw_q2low
cat("\nWilcoxon-Mann-Whitney worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
wmw_q2high
#cat("\nWilcoxon-Mann-Whitney worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.9\n")
#wmw_q2veryhigh
cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
survivoronly_q2low
cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
survivoronly_q2high
#cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.9\n")
#survivoronly_q2veryhigh

cat("\ndifference between WSR and WMW, q2 = 0.6\n")
wsr_q2low - wmw_q2low
cat("\ndifference between WSR and WMW, q2 = 0.8\n")
wsr_q2high - wmw_q2high
#cat("\ndifference between WSR and WMW, q2 = 0.9\n")
#wsr_q2veryhigh - wmw_q2veryhigh

cat("\ndifference between WSR and survivor only, q2 = 0.6\n")
wsr_q2low - survivoronly_q2low
cat("\ndifference between WSR and survivor only, q2 = 0.8\n")
wsr_q2high - survivoronly_q2high
#cat("\ndifference between WSR and survivor only, q2 = 0.9\n")
#wsr_q2veryhigh - survivoronly_q2veryhigh
```


#### 이론적 Power 계산

* $\mu = Np_1 + {N(N-1) \over 2}p_2$,
* $\sigma^2 = Np_1(1-p_1) + {N(N-1) \over 2}p_2(1-p_2) + 2N(N-1)(p_3 - p_1p_2) + N(N-1)(N-2)(p_4 - p_2^2)$
  
* $p_1 = P(\tilde{Y_1} > 0)$
* $p_2 = P(\tilde{Y_1} + \tilde{Y_2} >0)$
* $p_3 = P(\tilde{Y_1} + \tilde{Y_2} >0, \tilde{Y_1} >0)$ 
* $p_4 = P(\tilde{Y_1} + \tilde{Y_2} >0, \tilde{Y_1} + \tilde{Y_3} >0)$
  
* power : $P(W < \mu_0 - \sigma_0 z_\alpha) \simeq \Phi({\mu_0 - \mu - z_\alpha\sigma_0 \over \sigma})$
  
  
  
```{r}
n_prob = 1000000
z_alpha = qnorm(1 - alpha)
mu_0 = n * (n+1) / 4
sigma_squared_0 = n * (n+1) * (2*n+1) / 24

simulate_probs_exp <- function(n, q2, HR, Delta, time, rho_X, rho_t){
  lambda2 = -1 / time * log(q2)
  lambda1 = lambda2 * HR
  #q1 = exp(-lambda1 * t)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvexp(n, rate=c(lambda1, lambda2), corr=corr_t)
  t1 = t[,1]
  t2 = t[,2]
  
  corr_X = matrix(c(1, rho_X, rho_X, 1), nrow = 2)
  X = rmvnorm(n, mean = c(0, 2^0.5 * Delta), cov = corr_X)
  X1 = X[,1]
  X2 = X[,2]
  
  eta = min(X) - 1 - time - 2 *(max(X)-min(X)) # in order to be worst
  distmin = min(abs(X1-X2))/time/10 # less than minimum difference between X1 and X2 divided by time
  
  X1_tilde = X1 * (t1 >= time) + (eta + t1*distmin) * (t1 < time)
  X2_tilde = X2 * (t2 >= time) + (eta + t2*distmin) * (t2 < time)
  
  Y_tilde = X1_tilde - X2_tilde
  Y_tilde2 = permute(Y_tilde)
  Y_tilde3 = permute(Y_tilde)
  
  p1 = mean(Y_tilde > 0)
  p2 = mean((Y_tilde + Y_tilde2) > 0)
  p3 = mean(Y_tilde > 0 & ((Y_tilde + Y_tilde2) > 0)) # p3 = (p1 ** 2 + p2) / 2
  p4 = mean(((Y_tilde + Y_tilde2) > 0) * ((Y_tilde + Y_tilde3) > 0))
  return (c(p1,p2,p3,p4))
}

power.theory.array = array(data = NA, dim = c(2, length(q2), length(Delta), length(HR)), 
                           dimnames = list(c("Worst rank", "Survivors-only"), c("q2=60%", "q2=80%", "q2=90%"), 
                                           c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6), 
                                           c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)))


for (i in 1:length(q2)) {
  cat("q2[i] : ", q2[i], "\n")
  for (j in 1:length(Delta)) {
    cat("Delta[j] : ", Delta[j], "\n")
    for (k in 1:length(HR)) {
      probs = simulate_probs_exp(n_prob, q2[i], HR[k], Delta[j], time, rho_X, rho_t)
      mu = n * probs[1] + n * (n-1) / 2 * probs[2]
      sigma_squared = n * probs[1] * (1-probs[1]) + n * (n-1) / 2 * probs[2] * (1-probs[2]) + 2 * n * (n-1) * (probs[3] - probs[1] * probs[2]) + n * (n-1) * (n-2) * (probs[4] - probs[2]^2)
      
      power_theory = pnorm((-z_alpha * sqrt(sigma_squared_0)+ mu_0 - mu) / sqrt(sigma_squared))
      
      power.theory.array[1, i, j, k] = power_theory
      power.theory.array[2, i, j, k] = 0
    }
  }
}
```


```{r}
wsr_q2low_theory = power.theory.array[1,1,,]
wsr_q2high_theory = power.theory.array[1,2,,]
wsr_q2veryhigh_theory = power.theory.array[1,3,,]

#### comparison between simulation and theory power of wilcoxon signed worst-rank

cat("Wilcoxon signed worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.6\n")
wsr_q2low - wsr_q2low_theory
cat("Wilcoxon-Mann-Whitney worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.8\n")
wsr_q2high - wsr_q2high_theory
cat("Wilcoxon-Mann-Whitney worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.9\n")
wsr_q2veryhigh - wsr_q2veryhigh_theory
```



### (b) death follows weibull distribution

* $\begin{pmatrix} t_1 \\ t_2 \end{pmatrix} \sim {Weibull}\left(\begin{pmatrix} \lambda_{1}\\ \lambda_{2} \end{pmatrix}, \begin{pmatrix} k_{1}\\ k_{2} \end{pmatrix}, \rho_t \right), \quad \rho_t = 0.8, \quad k_1 = k_2 = 1.2$ 
* marginal pdf $f(t; \lambda, k) = \frac{k}{\lambda} \left(\frac{x}{\lambda}\right)^{k-1}e^{-\left(\frac{x}{\lambda}\right)^k} ,\quad t \ge 0$
* $q_i = e^{-\left(\frac{T}{\lambda_i}\right)^k}, \quad i=1,2, \quad T=3, \quad q_2 \in \{0.6, 0.8\}$
* $h_{i}(t) = \frac{f_{i}(t)}{S_{i}(t)} = \frac{kx^{k-1}}{\lambda^k}$
* $\text{HR} = h_{1}(t)/h_{2}(t) = (\lambda_2 / \lambda_1)^k$
  
  
#### generate data from `rmvweisd`
  
```{r}
Delta = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6)
q2 = c(0.6, 0.8)
shape = 1.2 #shape parameter
HR = c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)
time = 3
N = 1000
n = 50
rho_X = 0.8
rho_t = 0.8


simulate_weibull <- function(n, q2, HR, Delta, time, rho_X, rho_t, shape){
  lambda2 = time / (-log(q2))^(1/shape)
  lambda1 = lambda2 / HR^(1/shape)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvweibull(n, shape=shape, scale=c(lambda1, lambda2), corr=corr_t)
  t1 = t[,1]
  t2 = t[,2]
  
  corr_X = matrix(c(1, rho_X, rho_X, 1), nrow = 2)
  X = rmvnorm(n, mean = c(0, 2^0.5 * Delta), cov = corr_X)
  X1 = X[,1]
  X2 = X[,2]
  
  eta = min(X) - 1 - time - 2 *(max(X)-min(X)) # in order to be worst
  distmin = min(abs(X1-X2))/time/10 # less than minimum difference between X1 and X2 divided by time
  
  X1_tilde = X1 * (t1 >= time) + (eta + t1*distmin) * (t1 < time)
  X2_tilde = X2 * (t2 >= time) + (eta + t2*distmin) * (t2 < time)
  
  X1_survivorOnly = X1[t1 >= time & t2 >= time]
  X2_survivorOnly = X2[t1 >= time & t2 >= time]
  
  if (length(X1_survivorOnly) == 0) {
    return (c(-1,-1))
  }
  return (c(wilcox.test(X1_tilde, X2_tilde, alternative="less", paired=T)$p.value,
            wilcox.test(X1_survivorOnly, X2_survivorOnly, alternative="less", paired=T)$p.value))
}
```


#### Simulate

```{r}
power.weibull = array(data = NA, dim = c(2, length(q2), length(Delta), length(HR)), 
                      dimnames = list(c("Worst rank", "Survivors-only"), c("q2=60%", "q2=80%"), 
                                      c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6), 
                                      c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)))
set.seed(123)
for (i in 1:length(q2)) {
  for (j in 1:length(Delta)) {
    for (k in 1:length(HR)) {
      count = 0
      count_survivor = 0
      total = 0
      for (l in c(1:N)){
        wmw_tests = simulate_weibull(n, q2[i], HR[k], Delta[j], time, rho_X, rho_t, shape)
        wmw_p = wmw_tests[1]
        wmw_survivor_p = wmw_tests[2]
        if (wmw_p == -1) break;
        if (wmw_p <= 0.05){
          count = count + 1
        }
        if (wmw_survivor_p <= 0.05){
          count_survivor = count_survivor + 1
        }
        total = total + 1
      }
      power.weibull[1, i, j, k] = round(count/total, 3)
      power.weibull[2, i, j, k] = round(count_survivor/total, 3)
    }
  }
}
```

#### Result

```{r}
worstrank_q2low = power.weibull[1,1,,]
worstrank_q2high = power.weibull[1,2,,]
survivoronly_q2low = power.weibull[2,1,,]
survivoronly_q2high = power.weibull[2,2,,]

cat("worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
worstrank_q2low
cat("\nworst-rank simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
worstrank_q2high
cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
survivoronly_q2low
cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
survivoronly_q2high

cat("\ndifference between WSR and survivor only, q2 = 0.6\n")
worstrank_q2low - survivoronly_q2low
cat("\ndifference between WSR and survivor only, q2 = 0.8\n")
worstrank_q2high - survivoronly_q2high
```



#### ???론적 Power 계산

* $\mu = Np_1 + {N(N-1) \over 2}p_2$,
* $\sigma^2 = Np_1(1-p_1) + {N(N-1) \over 2}p_2(1-p_2) + 2N(N-1)(p_3 - p_1p_2) + N(N-1)(N-2)(p_4 - p_2^2)$
  
  * $p_1 = P(\tilde{Y_1} > 0)$
  * $p_2 = P(\tilde{Y_1} + \tilde{Y_2} >0)$
  * $p_3 = P(\tilde{Y_1} + \tilde{Y_2} >0, \tilde{Y_1} >0)$ 
  * $p_4 = P(\tilde{Y_1} + \tilde{Y_2} >0, \tilde{Y_1} + \tilde{Y_3} >0)$
  
  * power : $P(W < \mu_0 - \sigma_0 z_\alpha) \simeq \Phi({\mu_0 - \mu - z_\alpha\sigma_0 \over \sigma})$
  
  
  
```{r}
n_prob =1000000
z_alpha = qnorm(1 - alpha)
mu_0 = n * (n+1) / 4
sigma_squared_0 = n * (n+1) * (2*n+1) / 24

simulate_probs_weilbull <- function(n, q2, OR, Delta, time, rho_X, rho_t, shape){
  lambda2 = time / (-log(q2))^(1/shape)
  lambda1 = lambda2 / HR^(1/shape)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvweibull(n, shape=shape, scale=c(lambda1, lambda2), corr=corr_t)
  t1 = t[,1]
  t2 = t[,2]
  
  corr_X = matrix(c(1, rho_X, rho_X, 1), nrow = 2)
  X = rmvnorm(n, mean = c(0, 2^0.5 * Delta), cov = corr_X)
  X1 = X[,1]
  X2 = X[,2]
  
  eta = min(X) - 1 - time - 2 *(max(X)-min(X)) # in order to be worst
  distmin = min(abs(X1-X2))/time/10 # less than minimum difference between X1 and X2 divided by time
  
  X1_tilde = X1 * (t1 >= time) + (eta + t1*distmin) * (t1 < time)
  X2_tilde = X2 * (t2 >= time) + (eta + t2*distmin) * (t2 < time)
  
  Y_tilde = X1_tilde - X2_tilde
  Y_tilde2 = permute(Y_tilde)
  Y_tilde3 = permute(Y_tilde)
  
  p1 = mean(Y_tilde > 0)
  p2 = mean((Y_tilde + Y_tilde2) > 0)
  p3 = (p1**2 + p2) / 2
  p4 = mean(((Y_tilde + Y_tilde2) > 0) * ((Y_tilde + Y_tilde3) > 0))
  
  
  return (c(p1,p2,p3,p4))
}

power.weibull.theory = array(data = NA, dim = c(2, length(q2), length(Delta), length(HR)), 
                             dimnames = list(c("Worst rank", "Survivors-only"), c("q2=60%", "q2=80%"), 
                                             c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6), 
                                             c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)))


for (i in 1:length(q2)) {
  cat("q2[i] : ", q2[i], "\n")
  for (j in 1:length(Delta)) {
    cat("Delta[j] : ", Delta[j], "\n")
    for (k in 1:length(HR)) {
      probs = simulate_probs_weilbull(n_prob, q2[i], HR[k], Delta[j], time, rho_X, rho_t, shape)
      mu = n * probs[1] + n * (n-1) / 2 * probs[2]
      sigma_squared = n * probs[1] * (1-probs[1]) + n * (n-1) / 2 * probs[2] * (1-probs[2]) + 2 * n * (n-1) * (probs[3] - probs[1] * probs[2]) + n * (n-1) * (n-2) * (probs[4] - probs[2]^2)
      
      power_theory = pnorm((-z_alpha * sqrt(sigma_squared_0)+ mu_0 - mu) / sqrt(sigma_squared))
      
      power.weibull.theory[1, i, j, k] = power_theory
      power.weibull.theory[2, i, j, k] = 0
    }
  }
}
```


```{r}
wsr_q2low_theory = power.weibull.theory[1,1,,]
wsr_q2high_theory = power.weibull.theory[1,2,,]
```

#### comparison between simulation and theory power of wilcoxon signed worst-rank

```{r}
cat("Wilcoxon signed worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.6\n")
wsr_q2low - wsr_q2low_theory
mean(abs(wsr_q2low - wsr_q2low_theory))
cat("Wilcoxon-Mann-Whitney worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.8\n")
wsr_q2high - wsr_q2high_theory
mean(abs(wsr_q2high - wsr_q2high_theory))
```



### (d) Using Exp, With discrete survival times

#### function that generate data from `rmvexp` and define parameters

```{r}
#Define parameters 
Delta = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6)
q2 = c(0.6, 0.8)
HR = c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)
time = 3
N = 3000
n = 50
rho_X = 0.8
rho_t = 0.8
alpha = 0.05

simulate_exp <- function(n, q2, HR, Delta, time, rho_X, rho_t){
  lambda2 = -1 / time * log(q2)
  lambda1 = lambda2 * HR
  #q1 = exp(-lambda1 * t)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvexp(n, rate=c(lambda1, lambda2), corr=corr_t)
  t1 = ceiling(t[,1])
  t2 = ceiling(t[,2])
  
  corr_X = matrix(c(1, rho_X, rho_X, 1), nrow = 2)
  X = rmvnorm(n, mean = c(0, 2^0.5 * Delta), cov = corr_X)
  X1 = X[,1]
  X2 = X[,2]
  
  eta = min(X) - 1 - time - 2 *(max(X)-min(X)) # in order to be worst
  distmin = min(abs(X1-X2))/time/10 # less than minimum difference between X1 and X2 divided by time
  
  X1_tilde = X1 * (t1 > time) + (eta + t1*distmin) * (t1 <= time)
  X2_tilde = X2 * (t2 > time) + (eta + t2*distmin) * (t2 <= time)
  
  X1_survivorOnly = X1[t1 > time & t2 > time]
  X2_survivorOnly = X2[t1 > time & t2 > time]
  
  if (length(X1_survivorOnly) == 0) {
    return (c(-1,-1,-1))
  }
  return (c(wilcox.test(X1_tilde, X2_tilde, alternative="less", paired=T)$p.value,
            wilcox.test(X1_tilde, X2_tilde, alternative="less", paired=F)$p.value,
            wilcox.test(X1_survivorOnly, X2_survivorOnly, alternative="less", paired=T)$p.value))
}
```

#### Simulate(N = 3000, n = 50)

```{r}
power.exp = array(data = NA, dim = c(3, length(q2), length(Delta), length(HR)), 
                  dimnames = list(c("Paired Worst rank", "Non-paired Worst rank", "Survivors-only"),
                                  c("q2=60%", "q2=80%"), 
                                  c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6), 
                                  c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)))
set.seed(123)
for (i in 1:length(q2)) {
  for (j in 1:length(Delta)) {
    print(Delta[j])
    for (k in 1:length(HR)) {
      count_wsr=0
      count_wmw=0
      count_survivor = 0
      total = 0
      for (l in c(1:N)){
        wsr_tests = simulate_exp(n, q2[i], HR[k], Delta[j], time, rho_X, rho_t)
        wsr_p = wsr_tests[1]
        wmw_p = wsr_tests[2]
        wsr_survivor_p = wsr_tests[3]
        if (wsr_p == -1) break;
        if (wsr_p <= 0.05){
          count_wsr = count_wsr + 1
        }
        if (wmw_p <= 0.05){
          count_wmw = count_wmw +1
        }
        if (wsr_survivor_p <= 0.05){
          count_survivor = count_survivor + 1
        }
        total = total + 1
      }
      power.exp[1, i, j, k] = round(count_wsr/total, 3)
      power.exp[2, i, j, k] = round(count_wmw/total, 3)
      power.exp[3, i, j, k] = round(count_survivor/total, 3)
    }
  }
}
```

#### Result

```{r}
wsr_q2low = power.exp[1,1,,]
wsr_q2high = power.exp[1,2,,]

wmw_q2low = power.exp[2,1,,]
wmw_q2high = power.exp[2,2,,]

survivoronly_q2low = power.exp[3,1,,]
survivoronly_q2high = power.exp[3,2,,]

cat("Wilcoxon signed worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
wsr_q2low
cat("\nWilcoxon signed worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
wsr_q2high
cat("\nWilcoxon-Mann-Whitney worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
wmw_q2low
cat("\nWilcoxon-Mann-Whitney worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
wmw_q2high
cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
survivoronly_q2low
cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
survivoronly_q2high

cat("\ndifference between WSR and WMW, q2 = 0.6\n")
wsr_q2low - wmw_q2low
cat("\ndifference between WSR and WMW, q2 = 0.8\n")
wsr_q2high - wmw_q2high

cat("\ndifference between WSR and survivor only, q2 = 0.6\n")
wsr_q2low - survivoronly_q2low
cat("\ndifference between WSR and survivor only, q2 = 0.8\n")
wsr_q2high - survivoronly_q2high
```


#### 이론적 Power 계산

* $\mu = Np_1 + {N(N-1) \over 2}p_2$,
* $\sigma^2 = Np_1(1-p_1) + {N(N-1) \over 2}p_2(1-p_2) + 2N(N-1)(p_3 - p_1p_2) + N(N-1)(N-2)(p_4 - p_2^2)$
  
  * $p_1 = P(\tilde{Y_1} > 0)$
  * $p_2 = P(\tilde{Y_1} + \tilde{Y_2} >0)$
  * $p_3 = P(\tilde{Y_1} + \tilde{Y_2} >0, \tilde{Y_1} >0)$ 
  * $p_4 = P(\tilde{Y_1} + \tilde{Y_2} >0, \tilde{Y_1} + \tilde{Y_3} >0)$
  
  * power : $P(W < \mu_0 - \sigma_0 z_\alpha) \simeq \Phi({\mu_0 - \mu - z_\alpha\sigma_0 \over \sigma})$
  
  
  
```{r}
n_prob = 1000000
z_alpha = qnorm(1 - alpha)
mu_0 = n * (n+1) / 4
sigma_squared_0 = n * (n+1) * (2*n+1) / 24

simulate_probs_exp <- function(n, q2, HR, Delta, time, rho_X, rho_t){
  lambda2 = -1 / time * log(q2)
  lambda1 = lambda2 * HR
  #q1 = exp(-lambda1 * t)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvexp(n, rate=c(lambda1, lambda2), corr=corr_t)
  t1 = ceiling(t[,1])
  t2 = ceiling(t[,2])
  
  corr_X = matrix(c(1, rho_X, rho_X, 1), nrow = 2)
  X = rmvnorm(n, mean = c(0, 2^0.5 * Delta), cov = corr_X)
  X1 = X[,1]
  X2 = X[,2]
  
  eta = min(X) - 1 - time - 2 *(max(X)-min(X)) # in order to be worst
  distmin = min(abs(X1-X2))/time/10 # less than minimum difference between X1 and X2 divided by time
  
  X1_tilde = X1 * (t1 > time) + (eta + t1*distmin) * (t1 <= time)
  X2_tilde = X2 * (t2 > time) + (eta + t2*distmin) * (t2 <= time)
  
  Y_tilde = X1_tilde - X2_tilde
  Y_tilde2 = permute(Y_tilde)
  Y_tilde3 = permute(Y_tilde)
  
  p1 = mean(Y_tilde > 0)
  p2 = mean((Y_tilde + Y_tilde2) > 0)
  p3 = mean(Y_tilde > 0 & ((Y_tilde + Y_tilde2) > 0)) # p3 = (p1 ** 2 + p2) / 2
  p4 = mean(((Y_tilde + Y_tilde2) > 0) * ((Y_tilde + Y_tilde3) > 0))
  return (c(p1,p2,p3,p4))
}

power.theory.array = array(data = NA, dim = c(2, length(q2), length(Delta), length(HR)), 
                           dimnames = list(c("Worst rank", "Survivors-only"), c("q2=60%", "q2=80%"), 
                                           c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6), 
                                           c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)))


for (i in 1:length(q2)) {
  cat("q2[i] : ", q2[i], "\n")
  for (j in 1:length(Delta)) {
    cat("Delta[j] : ", Delta[j], "\n")
    for (k in 1:length(HR)) {
      probs = simulate_probs_exp(n_prob, q2[i], HR[k], Delta[j], time, rho_X, rho_t)
      mu = n * probs[1] + n * (n-1) / 2 * probs[2]
      sigma_squared = n * probs[1] * (1-probs[1]) + n * (n-1) / 2 * probs[2] * (1-probs[2]) + 2 * n * (n-1) * (probs[3] - probs[1] * probs[2]) + n * (n-1) * (n-2) * (probs[4] - probs[2]^2)
      
      power_theory = pnorm((-z_alpha * sqrt(sigma_squared_0)+ mu_0 - mu) / sqrt(sigma_squared))
      
      power.theory.array[1, i, j, k] = power_theory
      power.theory.array[2, i, j, k] = 0
    }
  }
}
```


```{r}
wsr_q2low_theory = power.theory.array[1,1,,]
wsr_q2high_theory = power.theory.array[1,2,,]
```

#### comparison between simulation and theory power of wilcoxon signed worst-rank

```{r}
cat("Wilcoxon signed worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.6\n")
wsr_q2low - wsr_q2low_theory
cat("Wilcoxon-Mann-Whitney worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.8\n")
wsr_q2high - wsr_q2high_theory
```