---
title: "WSR_power_analysis_exp"
author: "민태원, 권대영"
date: "`r Sys.Date()`"
output: html_document
---
  
#### load `lcmix` package which includes `rmvnorm` and `rmvexp`.

* simulate_exp generates 3 p-values at once: worst-rank wsr, worst-rank wmw, survivors only wsr
  
```{r, include=F}
library(combinat)
library(flexsurv)
library(gtools)
source("C:/Users/MIN/OneDrive - SNU/문서/GitHub/Matching_WorstRank/Simulation/mvfunctions.R")
```


## (1) Specification

* Number of simulations = 5000
* one-sided, alpha = 0.05
* $\Delta_x = \frac{\mu_{x2} - \mu_{x1}}{\sqrt{2}\sigma_{x1}} \in \{0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6\}$ is the primary outcome of interest ($\mu_{x1} = 0$, $\sigma_{x1} = 1$)
* $\begin{pmatrix} X_1 \\ X_2 \end{pmatrix} \sim N \begin{pmatrix} \begin{pmatrix} 0\\\sqrt{2} \Delta_x \end{pmatrix}, \begin{pmatrix} \sigma^2_{x1} & \rho_x\sigma_{x1}\sigma_{x2} \\ \rho_x\sigma_{x1}\sigma_{x2} & \sigma^2_{x2} \end{pmatrix} \end{pmatrix}$, $\rho_x=0.8$, $\sigma_{x1}=\sigma_{x2}=1$
  * $p_i = E(\delta_{ij}) = \mathrm{Pr}(\delta_{ij}=1) = \mathrm{Pr}(t_{ij} \leq T)$ is the probability of truncated by death at observed time T for group i
* $q_i = 1-p_i = \mathrm{Pr}(t_{ij} > T), \quad i = 1, 2$ is the probability of survival at observed time T
* $q_2 \in \{0.6, 0.8, 0.9\}, \quad T = 3$
* $\text{HR (Hazard Ratio)} \in \{1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0\}$
  
  
  
  
## (2) Power computation by Simulation

### (a) death follows exponential distribution

* $\begin{pmatrix} t_1 \\ t_2 \end{pmatrix} \sim Exp\left( \begin{pmatrix} \lambda_{1}\\ \lambda_{2} \end{pmatrix}, \rho_t \right), \quad \rho_t = 0.8$ 
* marginal pdf $f(t; \lambda) = \lambda e^{-\lambda t} ,\quad t \ge 0$
* $q_i = exp(-\lambda_i T), \quad i=1,2, \quad T=3, \quad q_2 \in \{0.6, 0.8, 0.9\}$
* $h_{i}(t) = \lambda_i$
* $\text{HR} = h_{1}(t)/h_{2}(t) = \lambda_1 / \lambda_2$

#### define parameters
  
```{r}
#Define parameters 
Delta = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6)
q2 = c(0.6, 0.8, 0.9)
HR = c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)
time = 3
N = 5000
n = 50
rho_X = 0.8
rho_t = 0.8
alpha = 0.05

power.exp = array(data = NA, dim = c(3, length(q2), length(Delta), length(HR)), 
                  dimnames = list(c("Paired Worst rank", "Non-paired Worst rank", "Survivors-only"),
                                  c("q2=60%", "q2=80%", "q2=90%"), 
                                  c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6), 
                                  c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)))
```

#### Simulate(N = 5000, n = 50)

```{r, include = F, echo = F}
set.seed(123)
for (i in 1:length(q2)) {
  for (j in 1:length(Delta)) {
    for (k in 1:length(HR)) {
      cat("q2 value: ",q2[i], ", ")
      cat("Delta value: ",Delta[j], ", ")
      cat("HR value: ",HR[k], "\n")
      count_wsr=0
      count_wmw=0
      count_survivor = 0
      total = 0
      for (l in c(1:N)){
        wsr_tests = simulate_exp(n, q2[i], HR[k], Delta[j], time, rho_X, rho_t)
        wsr_p = wsr_tests[1]
        wmw_p = wsr_tests[2]
        wsr_survivor_p = wsr_tests[3]
        if (wsr_p == -1) break;
        if (wsr_p <= 0.05){
          count_wsr = count_wsr + 1
        }
        if (wmw_p <= 0.05){
          count_wmw = count_wmw +1
        }
        if (wsr_survivor_p <= 0.05){
          count_survivor = count_survivor + 1
        }
        total = total + 1
      }
      power.exp[1, i, j, k] = round(count_wsr/total, 3)
      power.exp[2, i, j, k] = round(count_wmw/total, 3)
      power.exp[3, i, j, k] = round(count_survivor/total, 3)
    }
  }
}
```

#### Result

```{r}
wsr_q2low = power.exp[1,1,,]
wsr_q2mid = power.exp[1,2,,]
wsr_q2high = power.exp[1,3,,]

wmw_q2low = power.exp[2,1,,]
wmw_q2mid = power.exp[2,2,,]
wmw_q2high = power.exp[2,3,,]

survivoronly_q2low = power.exp[3,1,,]
survivoronly_q2mid = power.exp[3,2,,]
survivoronly_q2high = power.exp[3,3,,]

diffwsrsur_q2low = wsr_q2low - survivoronly_q2low
diffwsrsur_q2mid = wsr_q2mid - survivoronly_q2mid
diffwsrsur_q2high = wsr_q2high - survivoronly_q2high

cat("Wilcoxon signed worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
wsr_q2low
cat("\nWilcoxon signed worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
wsr_q2mid
cat("\nWilcoxon signed worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.9\n")
wsr_q2high
cat("\nWilcoxon-Mann-Whitney worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
wmw_q2low
cat("\nWilcoxon-Mann-Whitney worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
wmw_q2mid
cat("\nWilcoxon-Mann-Whitney worst-rank simulation power, rows are Delta, cols are HR, q2 = 0.9\n")
wmw_q2high
cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
survivoronly_q2low
cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
survivoronly_q2mid
cat("\nsurvivors only simulation power, rows are Delta, cols are HR, q2 = 0.9\n")
survivoronly_q2high

cat("\ndifference between WSR and survivor only, q2 = 0.6\n")
diffwsrsur_q2low
cat("\ndifference between WSR and survivor only, q2 = 0.8\n")
diffwsrsur_q2mid
cat("\ndifference between WSR and survivor only, q2 = 0.9\n")
diffwsrsur_q2high
```


### 이론적 Power 계산

* $\mu = Np_1 + {N(N-1) \over 2}p_2$,
* $\sigma^2 = Np_1(1-p_1) + {N(N-1) \over 2}p_2(1-p_2) + 2N(N-1)(p_3 - p_1p_2) + N(N-1)(N-2)(p_4 - p_2^2)$
  
* $p_1 = P(\tilde{Y_1} > 0)$
* $p_2 = P(\tilde{Y_1} + \tilde{Y_2} >0)$
* $p_3 = P(\tilde{Y_1} + \tilde{Y_2} >0, \tilde{Y_1} >0)$ 
* $p_4 = P(\tilde{Y_1} + \tilde{Y_2} >0, \tilde{Y_1} + \tilde{Y_3} >0)$
  
* power : $P(W < \mu_0 - \sigma_0 z_\alpha) \simeq \Phi({\mu_0 - \mu - z_\alpha\sigma_0 \over \sigma})$
  
* same setting as before
  
```{r}

Delta = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6)
q2 = c(0.6, 0.8, 0.9)
HR = c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)
time = 3
N = 5000
n = 50
rho_X = 0.8
rho_t = 0.8
alpha = 0.05

power.theory.array = array(data = NA, dim = c(length(q2), length(Delta), length(HR)), 
                           dimnames = list(c("q2=60%", "q2=80%", "q2=90%"), 
                                           c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6), 
                                           c(1.0, 1.2, 1.4, 1.6, 2.0, 2.4, 3.0)))

n_prob = 1000000
z_alpha = qnorm(1 - alpha)
mu_0 = n * (n+1) / 4
sigma_squared_0 = n * (n+1) * (2*n+1) / 24
```
  
  
```{r, include = F, echo = F}
for (i in 1:length(q2)) {
  for (j in 1:length(Delta)) {
    for (k in 1:length(HR)) {
      cat("q2[i] : ", q2[i], ", ")
      cat("Delta[j] : ", Delta[j], ", ")
      cat("HR: ", HR[k], "\n")
      probs = simulate_probs_exp(n_prob, q2[i], HR[k], Delta[j], time, rho_X, rho_t)
      mu = n * probs[1] + n * (n-1) / 2 * probs[2]
      sigma_squared = n * probs[1] * (1-probs[1]) + n * (n-1) / 2 * probs[2] * (1-probs[2]) + 2 * n * (n-1) * (probs[3] - probs[1] * probs[2]) + n * (n-1) * (n-2) * (probs[4] - probs[2]^2)
      power_theory = pnorm((-z_alpha * sqrt(sigma_squared_0)+ mu_0 - mu) / sqrt(sigma_squared))
      
      power.theory.array[i, j, k] = power_theory
    }
  }
}
```


```{r}
wsr_q2low_theory = power.theory.array[1,,]
wsr_q2mid_theory = power.theory.array[2,,]
wsr_q2high_theory = power.theory.array[3,,]

difftheory_q2low = wsr_q2low - wsr_q2low_theory
difftheory_q2mid = wsr_q2mid - wsr_q2mid_theory
difftheory_q2high = wsr_q2high - wsr_q2high_theory

#### comparison between simulation and theory power of wilcoxon signed worst-rank

cat("Wilcoxon signed worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.6\n")
difftheory_q2low
cat("Wilcoxon signed worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.8\n")
difftheory_q2mid
cat("Wilcoxon signed worst-rank simulation power - theory power, rows are Delta, cols are HR, q2 = 0.9\n")
difftheory_q2high
```

#### save to csv

```{r, include = F, echo = F}
write.csv(wsr_q2low, "tables/wsr_q2low.csv")
write.csv(wsr_q2mid, "tables/wsr_q2mid.csv")
write.csv(wsr_q2high, "tables/wsr_q2high.csv")

write.csv(wmw_q2low, "tables/wmw_q2low.csv")
write.csv(wmw_q2mid, "tables/wmw_q2mid.csv")
write.csv(wmw_q2high, "tables/wmw_q2high.csv")

write.csv(survivoronly_q2low, "tables/survivoronly_q2low.csv")
write.csv(survivoronly_q2mid, "tables/survivoronly_q2mid.csv")
write.csv(survivoronly_q2high, "tables/survivoronly_q2high.csv")

write.csv(diffwsrsur_q2low, "tables/diffwsrsur_q2low.csv")
write.csv(diffwsrsur_q2mid, "tables/diffwsrsur_q2mid.csv")
write.csv(diffwsrsur_q2high, "tables/diffwsrsur_q2high.csv")

write.csv(difftheory_q2low, "tables/difftheory_q2low.csv")
write.csv(difftheory_q2mid, "tables/difftheory_q2mid.csv")
write.csv(difftheory_q2high, "tables/difftheory_q2high.csv")
```




