---
title: "Power Comparison between Sequential and truncated method_20230115"
author: "민태원, 권대영"
date: "`r Sys.Date()`"
output: html_document
---

```{r include=F}
#install.packages("nnls") #install before download lcmix
#install.packages("lcmix", repos="http://R-Forge.R-project.org")
#install.packages("flexsurv")
#library(lcmix)
library(combinat)
library(flexsurv)
library(sensitivitymv)
library(gtools)
source("mvfunctions.R")
```

## (1) Specification

* N: Number of simulations = 5000
* n: Number of generated samples per group = 500
* Group2 is treated group 
* one-sided, alpha = 0.05, beta = 0.025
* $\Delta_x = \frac{\mu_{x2} - \mu_{x1}}{\sqrt{2}\sigma_{x1}} \in \{0.0, 0.05, 0.1, 0.15, 0.20\}$ is the primary outcome of interest ($\mu_{x1} = 0$, $\sigma_{x1} = 1$)
* $\begin{pmatrix} X_1 \\ X_2 \end{pmatrix} \sim N \begin{pmatrix} \begin{pmatrix} 0\\\sqrt{2} \Delta_x \end{pmatrix}, \begin{pmatrix} \sigma^2_{x1} & \rho_x\sigma_{x1}\sigma_{x2} \\ \rho_x\sigma_{x1}\sigma_{x2} & \sigma^2_{x2} \end{pmatrix} \end{pmatrix}$, $\rho_x=0.8$, $\sigma_{x1}=\sigma_{x2}=1$
* $p_i = E(\delta_{ij}) = \mathrm{Pr}(\delta_{ij}=1) = \mathrm{Pr}(t_{ij} \leq T)$ is the probability of truncated by death at observed time T for group i
* $q_i = 1-p_i = \mathrm{Pr}(t_{ij} > T), \quad i = 1, 2$ is the probability of survival at observed time T
* $q_2 \in \{0.6, 0.8, 0.9\}, \quad T = 3$
* $\text{HR (Hazard Ratio)} \in \{1.0, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30  \}$

## (2) Power Simulation

### (a) Survival time follows exponential distribution

#### Power for sequential test, used 0.025, 0.025 as alpha


```{r}
#Define parameters 
Delta = c(0.0, 0.05, 0.1, 0.15, 0.20)
q2 = c(0.6, 0.8, 0.9)
HR = c(1.0, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30)
time = 3
N = 5000
n = 500
rho_t = 0.8
rho_X = 0.8
alpha = 0.05
beta = 0.025

simulate.exp.binom <- function(n, q2, HR, time, rho_t){
  lambda2 = -1 / time * log(q2)
  lambda1 = lambda2 * HR
  #q1 = exp(-lambda1 * t)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvexp(n, rate=c(lambda1, lambda2), corr=corr_t)
  t1 = t[,1]
  t2 = t[,2]
  
  OnlyControlDead = sum(t1 < time & t2 >= time)
  OnlyTreatedDead = sum(t2 < time & t1 >= time)
  if (OnlyControlDead == 0 & OnlyTreatedDead == 0) return (-1); # no discordant pair
  pval = binom.test(c(OnlyControlDead, OnlyTreatedDead), alternative = "greater")$p.value;
  if (pval <= alpha) {return (pval);}
  else { 
    if (OnlyTreatedDead == 0) {pval.alt = 0;}
    else {pval.alt = binom.test(c(OnlyControlDead+1, OnlyTreatedDead-1), alternative = "greater")$p.value;}
    
    if (pval.alt > alpha) { return (pval); }
    else {
      #cat("prob: ", (alpha - pval.alt)/(pval - pval.alt), "pval_alt: ", pval.alt, "\n");
      return (1 - rbinom(1, 1, (alpha - pval.alt)/(pval - pval.alt))); # auxiliary random
    }
  }
}

simulate.exp.both <- function(n, q2, HR, Delta, time, rho_X, rho_t){
  lambda2 = -1 / time * log(q2)
  lambda1 = lambda2 * HR
  #q1 = exp(-lambda1 * t)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvexp(n, rate=c(lambda1, lambda2), corr=corr_t)
  t1 = t[,1]
  t2 = t[,2]
  
  corr_X = matrix(c(1, rho_X, rho_X, 1), nrow = 2)
  X = rmvnorm(n, mean = c(0, 2^0.5 * Delta), cov = corr_X)
  X1 = X[,1]
  X2 = X[,2]
  
  eta = min(X) - 1 - time - 2 *(max(X)-min(X)) # in order to be worst
  distmin = min(abs(X1-X2))/time/10 # less than minimum difference between X1 and X2 divided by time
  
  X1_tilde = X1 * (t1 >= time) + (eta + t1*distmin) * (t1 < time)
  X2_tilde = X2 * (t2 >= time) + (eta + t2*distmin) * (t2 < time)
  
  X1_survivorOnly = X1[t1 >= time & t2 >= time]
  X2_survivorOnly = X2[t1 >= time & t2 >= time]
  
  if (length(X1_survivorOnly) == 0) {
    return (c(-1,-1,-1))
  }
  
  pval.2 = wilcox.test(X1_tilde, X2_tilde, alternative="less", paired=T)$p.value
  
  OnlyControlDead = sum(t1 < time & t2 >= time)
  OnlyTreatedDead = sum(t2 < time & t1 >= time)
  if (OnlyControlDead == 0 & OnlyTreatedDead == 0) return (c(-1,-1)); # no discordant pair
  pval = binom.test(c(OnlyControlDead, OnlyTreatedDead), alternative = "greater")$p.value;
  if (pval <= alpha) {return (c(pval, pval.2));}
  else { 
    if (OnlyTreatedDead == 0) {pval.alt = 0;}
    else {pval.alt = binom.test(c(OnlyControlDead+1, OnlyTreatedDead-1), alternative = "greater")$p.value;}
    
    if (pval.alt > alpha) { return (c(pval, pval.2)); }
    else {
      #cat("prob: ", (alpha - pval.alt)/(pval - pval.alt), "pval_alt: ", pval.alt, "\n");
      return (c(1 - rbinom(1, 1, (alpha - pval.alt)/(pval - pval.alt)), pval.2)); # auxiliary random, 
    }
  }
}
```

```{r}
power.exp.seq = array(data = NA, dim = c(length(q2), length(Delta), length(HR)), 
                    dimnames = list(c("q2=60%", "q2=80%", "q2=90%"), 
                                    c(0.0, 0.05, 0.10, 0.15, 0.20), 
                                    c(1.0, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30)))
set.seed(123)
for (i in 1:length(q2)) {
  print(q2[i])
  for (j in 1:length(Delta)) {
    print(Delta[j])
    for (k in 1:length(HR)) {
      count = 0
      total = 0
      for (l in c(1:N)){
        pvals = simulate.exp.both(n, q2[i], HR[k], Delta[j], time, rho_X, rho_t)
        if (pvals[1] == -1) {
          print(l)
          next;
        }  
        if (pvals[2] == -1) {
          print(l)
          next;
        }  
        if (pvals[1] <= beta){
          count = count + 1
        } else{
          if (pvals[2] <= (alpha - beta)){
            count = count + 1
          }
        }
        total = total + 1
      }
      power.exp.seq[i, j, k] = round(count/total, 3)
    }
  }
}
```


```{r}
exp.seq.q2low = power.exp.seq[1,,]
exp.seq.q2mid = power.exp.seq[2,,]
exp.seq.q2high = power.exp.seq[3,,]

cat("Survival Time Follows Expoential Distribution\n")

cat("sequential testing simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
exp.seq.q2low
cat("sequential testing simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
exp.seq.q2mid
cat("sequential testing simulation power, rows are Delta, cols are HR, q2 = 0.9\n")
exp.seq.q2high
```

#### Power for TruncatedP using death and secondary outcome

```{r}
#Define parameters 
Delta = c(0.0, 0.05, 0.1, 0.15, 0.20)
q2 = c(0.6, 0.8, 0.9)
HR = c(1.0, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30)
time = 3
N = 5000
n = 500
rho_t = 0.8
alpha = 0.05

power.exp.truncP = array(data = NA, dim = c(length(q2), length(Delta), length(HR)),
                    dimnames = list(c("q2=60%", "q2=80%", "q2=90%"),
                                    c(0.0, 0.05, 0.1, 0.15, 0.20),
                                    c(1.0, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30)))

set.seed(123)

pval.transform <- function(pvals, r){
  p1 = pvals[1]
  p2 = pvals[2]
  return (c(p1, 1 - pnorm((qnorm(1 - p2) - r * qnorm(1 - p1)) / sqrt(1 - r**2))))  
}

for (i in 1:length(q2)) {
  print(q2[i])
  for (j in 1:length(Delta)) {
    print(Delta[j])
    for (k in 1:length(HR)) {
      print(HR[k])
      count = 0
      total = 0
      pvals.forcov = array(data = NA, dim = c(N, 2))
      for (l in c(1:N)){
        pvals.forcov[l,] = simulate.exp.both(n, q2[i], HR[k], Delta[j], time, rho_X, rho_t)
      }
      r = cor(pvals.forcov)[1,2]
      cat("cor matrix: ", cor(pvals.forcov))
      cat("For q2: ", q2[i], ", Delta: ",Delta[j], ", HR: ",HR[k])
      for (l in c(1:N)){
        pvals.forcov[l,] = pval.transform(pvals.forcov[l,], r)
        if (!is.finite(pvals.forcov[l,1]) | !is.finite(pvals.forcov[l,2])){
          next;
        }
        if (pvals.forcov[l,1] == 0 | pvals.forcov[l,2] == 0){
          next;
        }
        pval = truncatedP(pvals.forcov[l,])
        if (pval <= alpha){
          count = count + 1
        }
        total = total + 1
      }
      cat(", total: ", total, "\n")
      power.exp.truncP[i, j, k] = round(count/total, 3)
    }
  }
}

```

```{r}
exp.truncP.q2low = power.exp.truncP[1,,]
exp.truncP.q2mid = power.exp.truncP[2,,]
exp.truncP.q2high = power.exp.truncP[3,,]

cat("Survival Time Follows Expoential Distribution\n")

cat("truncatedP simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
exp.truncP.q2low
cat("truncatedP simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
exp.truncP.q2mid
cat("truncatedP simulation power, rows are Delta, cols are HR, q2 = 0.9\n")
exp.truncP.q2high
```
```{r}
cat("Difference, (sequential testing) - (truncatedP method)\n")

cat("seq - truncP simulation power, rows are Delta, cols are HR, q2 = 0.6\n")
exp.seq.q2low - exp.truncP.q2low
cat("seq - truncP simulation power, rows are Delta, cols are HR, q2 = 0.8\n")
exp.seq.q2mid - exp.truncP.q2mid
cat("seq - truncP simulation power, rows are Delta, cols are HR, q2 = 0.9\n")
exp.seq.q2high - exp.truncP.q2high
```



### (b) Survival time follows weibull distribution

