---
title: "Power curves for 4 scenarios"
author: "권대영, 민태원"
date: "`r Sys.Date()`"
output: html_document
---

```{r include=F}
library(combinat)
library(flexsurv)
library(sensitivitymv)
library(gtools)
library(ggplot2)
source("mvfunctions.R")
```

## (1) Specification: Choose 4 scenarios (HR, Delta_y)

* N: Number of simulations = 5,000
* n: Number of generated samples per group = 500
* Group2 is treated group 
* one-sided, alpha = 0.05, beta = 0.025
* $\Delta_y = \frac{\mu_{y2} - \mu_{y1}}{\sqrt{2}\sigma_{y1}} \in \{0.0, 0.05\}$ is the primary outcome of interest ($\mu_{y1} = 0$, $\sigma_{y1} = 1$)
* $\begin{pmatrix} Y_1 \\ Y_2 \end{pmatrix} \sim N \begin{pmatrix} \begin{pmatrix} 0\\\sqrt{2} \Delta_y \end{pmatrix}, \begin{pmatrix} \sigma^2_{y1} & \rho_y\sigma_{y1}\sigma_{y2} \\ \rho_y\sigma_{y1}\sigma_{y2} & \sigma^2_{y2} \end{pmatrix} \end{pmatrix}$, $\rho_y=0.8$, $\sigma_{y1}=\sigma_{y2}=1$
* $p_i = E(\delta_{ij}) = \mathrm{Pr}(\delta_{ij}=1) = \mathrm{Pr}(t_{ij} \leq T)$ is the probability of truncated by death at observed time T for group i
* $q_i = 1-p_i = \mathrm{Pr}(t_{ij} > T), \quad i = 1, 2$ is the probability of survival at observed time T
* $q_2 \in \{0.6, 0.8, 0.9\}=\{low, mid, high\}, \quad T = 3$
* $\text{HR(Hazard Ratio)} \in \{1, 1.05\}$


## (2) Power Simulation

### (a) Survival time follows exponential distribution

#### Power for sequential test

```{r}
#Define parameters 
Delta = c(0.0, 0.05)
q2 = c(0.6, 0.8, 0.9)
HR = c(1.0, 1.05)
time = 3
N = 5000
n = 500
rho_t = 0.8
rho_Y = 0.8
alpha = 0.05
beta = seq(0, 0.05, by=5e-4)

simulate.exp.both <- function(n, q2, HR, Delta, time, rho_Y, rho_t){
  lambda2 = -1 / time * log(q2)
  lambda1 = lambda2 * HR
  #q1 = exp(-lambda1 * t)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvexp(n, rate=c(lambda1, lambda2), corr=corr_t)
  t1 = t[,1]
  t2 = t[,2]
  
  corr_Y = matrix(c(1, rho_Y, rho_Y, 1), nrow = 2)
  Y = rmvnorm(n, mean = c(0, 2^0.5 * Delta), cov = corr_Y)
  Y1 = Y[,1]
  Y2 = Y[,2]
  
  eta = min(Y) - 1 - time - 2 *(max(Y)-min(Y)) # in order to be worst
  distmin = min(abs(Y1-Y2))/time/10 # less than minimum difference between Y1 and Y2 divided by time
  
  Y1_tilde = Y1 * (t1 >= time) + (eta + t1*distmin) * (t1 < time)
  Y2_tilde = Y2 * (t2 >= time) + (eta + t2*distmin) * (t2 < time)
  
  Y1_survivorOnly = Y1[t1 >= time & t2 >= time]
  Y2_survivorOnly = Y2[t1 >= time & t2 >= time]
  
  if (length(Y1_survivorOnly) == 0) {
    return (c(-1,-1,-1))
  }
  
  pval.2 = wilcox.test(Y1_tilde, Y2_tilde, alternative="less", paired=T)$p.value
  
  OnlyControlDead = sum(t1 < time & t2 >= time)
  OnlyTreatedDead = sum(t2 < time & t1 >= time)
  if (OnlyControlDead == 0 & OnlyTreatedDead == 0) return (c(-1,-1)); # no discordant pair
  pval = binom.test(c(OnlyControlDead, OnlyTreatedDead), alternative = "greater")$p.value;
  if (pval <= alpha) {return (c(pval, pval.2));}
  else { 
    if (OnlyTreatedDead == 0) {pval.alt = 0;}
    else {pval.alt = binom.test(c(OnlyControlDead+1, OnlyTreatedDead-1), alternative = "greater")$p.value;}
    
    if (pval.alt > alpha) { return (c(pval, pval.2)); }
    else {
      return (c(1 - rbinom(1, 1, (alpha - pval.alt)/(pval - pval.alt)), pval.2)); # auxiliary random 
    }
  }
}
```


```{r}
power.exp.seq = array(data = NA, dim=c(length(q2), length(Delta), length(HR), length(beta)), dimnames = list(c("q2=60%", "q2=80%", "q2=90%"), c(0.0, 0.05), c(1.0, 1.05), seq(0, 0.05, by=5e-4)))

set.seed(123)
for (i in 1:length(q2)) {
  print(q2[i])
  for (j in 1:length(Delta)) {
    print(Delta[j])
    for (k in 1:length(HR)) {
      print(HR[k])
      count = array(0, length(beta))
      total = 0
      for (l in c(1:N)) {
        pvals = simulate.exp.both(n, q2[i], HR[k], Delta[j], time, rho_Y, rho_t)
        if (pvals[1] == -1 || pvals[2] == -1) {
          print(l)
          next;
        }
        total = total + 1
        for (b in 1:length(beta)) {
          if (pvals[1] <= beta[b]) {count[b] = count[b] + 1}
          else if (pvals[2] <= (alpha - beta[b])) {count[b] = count[b] + 1}
        }
      }
      power.exp.seq[i,j,k,] = count/total
    }
  }
}
```

#### plot

```{r}
exp.seq.q2low = power.exp.seq[1,,,]
exp.seq.q2mid = power.exp.seq[2,,,]
exp.seq.q2high = power.exp.seq[3,,,]
```

##### (1) q2 = low (60%)

```{r}
power11 = exp.seq.q2low[1,1,] # Delta = 0, HR = 1
power12 = exp.seq.q2low[1,2,] # Delta = 0, HR = 1.05
power21 = exp.seq.q2low[2,1,] # Delta = 0.05, HR = 1
power22 = exp.seq.q2low[2,2,] # Delta = 0.05, HR = 1.05

df = data.frame(beta = rep(beta, 4), group=rep(c("11","12","21","22"), each=length(beta)), power = c(power11, power12, power21, power22))

df$beta[which(df$power[df$group=="11"] == max(df$power[df$group=="11"]))]
df$beta[which(df$power[df$group=="12"] == max(df$power[df$group=="12"]))]
df$beta[which(df$power[df$group=="21"] == max(df$power[df$group=="21"]))]
df$beta[which(df$power[df$group=="22"] == max(df$power[df$group=="22"]))]
```

```{r}
ggplot(df, aes(beta, power, col=group)) +
  geom_line(size=1)
```



##### (2) q2 = mid (80%)

```{r}
power11 = exp.seq.q2mid[1,1,] # Delta = 0, HR = 1
power12 = exp.seq.q2mid[1,2,] # Delta = 0, HR = 1.05
power21 = exp.seq.q2mid[2,1,] # Delta = 0.05, HR = 1
power22 = exp.seq.q2mid[2,2,] # Delta = 0.05, HR = 1.05

df = data.frame(beta = rep(beta, 4), group=rep(c("11","12","21","22"), each=length(beta)), power = c(power11, power12, power21, power22))

df$beta[which(df$power[df$group=="11"] == max(df$power[df$group=="11"]))]
df$beta[which(df$power[df$group=="12"] == max(df$power[df$group=="12"]))]
df$beta[which(df$power[df$group=="21"] == max(df$power[df$group=="21"]))]
df$beta[which(df$power[df$group=="22"] == max(df$power[df$group=="22"]))]
```

```{r}
ggplot(df, aes(beta, power, col=group)) +
  geom_line(size=1)
```



##### (3) q2 = high (90%)

```{r}
power11 = exp.seq.q2high[1,1,] # Delta = 0, HR = 1
power12 = exp.seq.q2high[1,2,] # Delta = 0, HR = 1.05
power21 = exp.seq.q2high[2,1,] # Delta = 0.05, HR = 1
power22 = exp.seq.q2high[2,2,] # Delta = 0.05, HR = 1.05

df = data.frame(beta = rep(beta, 4), group=rep(c("11","12","21","22"), each=length(beta)), power = c(power11, power12, power21, power22))

df$beta[which(df$power[df$group=="11"] == max(df$power[df$group=="11"]))]
df$beta[which(df$power[df$group=="12"] == max(df$power[df$group=="12"]))]
df$beta[which(df$power[df$group=="21"] == max(df$power[df$group=="21"]))]
df$beta[which(df$power[df$group=="22"] == max(df$power[df$group=="22"]))]
```

```{r}
ggplot(df, aes(beta, power, col=group)) +
  geom_line(size=1)
```
