---
title: "Power curves for each HR"
author: "권대영, 민태원"
date: "`r Sys.Date()`"
output: html_document
---

```{r include=F}
library(combinat)
library(flexsurv)
library(sensitivitymv)
library(gtools)
library(ggplot2)
source("mvfunctions.R")
library(gridExtra)
library(dplyr)
```   

## (1) Specification: Choose 16 scenarios (HR, Delta_y)

* N: Number of simulations = 5,000
* n: Number of generated samples per group = 50
* Group2 is treated group 
* one-sided, alpha = 0.05, beta = 0 ~ 0.05 (by 0.001)
* $\Delta_y = \frac{\mu_{y2} - \mu_{y1}}{\sqrt{2}\sigma_{y1}} \in \{0.0, 0.1, 0.2, 0.3\}$ is the primary outcome of interest ($\mu_{y1} = 0$, $\sigma_{y1} = 1$)
* $\begin{pmatrix} Y_1 \\ Y_2 \end{pmatrix} \sim N \begin{pmatrix} \begin{pmatrix} 0\\\sqrt{2} \Delta_y \end{pmatrix}, \begin{pmatrix} \sigma^2_{y1} & \rho_y\sigma_{y1}\sigma_{y2} \\ \rho_y\sigma_{y1}\sigma_{y2} & \sigma^2_{y2} \end{pmatrix} \end{pmatrix}$, $\rho_y=0.8$, $\sigma_{y1}=\sigma_{y2}=1$
* $p_i = E(\delta_{ij}) = \mathrm{Pr}(\delta_{ij}=1) = \mathrm{Pr}(t_{ij} \leq T)$ is the probability of truncated by death at observed time T for group i
* $q_i = 1-p_i = \mathrm{Pr}(t_{ij} > T), \quad i = 1, 2$ is the probability of survival at observed time T
* $q_2 \in \{0.6, 0.8, 0.9\}=\{low, mid, high\}, \quad T = 3$
* $\text{HR(Hazard Ratio)} \in \{1.0, 1.2, 1.4, 1.6\}$


## (2) Power Simulation

### (a) Survival time follows exponential distribution

#### Power for sequential test

- T1 = binom
- T2 = worst-rank
- T3 = survivor-only

```{r}
#Define parameters 
Delta = c(0.0, 0.1, 0.2, 0.3)
q2 = c(0.6, 0.8, 0.9)
HR = c(1.0, 1.2, 1.4, 1.6)
time = 3
N = 5000
n = 50
rho_t = 0.8
rho_Y = 0.8
alpha = 0.05
beta = seq(0, 0.05, by=0.001)
```

```{r}
generate.outcomes.exp <- function(n, q2, HR, Delta, time, rho_Y, rho_t){
  lambda2 = -1 / time * log(q2)
  lambda1 = lambda2 * HR
  #q1 = exp(-lambda1 * t)
  
  corr_t = matrix(c(1, rho_t, rho_t, 1), nrow = 2)
  t = rmvexp(n, rate=c(lambda1, lambda2), corr=corr_t)
  t1 = t[,1]
  t2 = t[,2]
  
  corr_Y = matrix(c(1, rho_Y, rho_Y, 1), nrow = 2)
  Y = rmvnorm(n, mean = c(0, 2^0.5 * Delta), cov = corr_Y)
  Y1 = Y[,1]
  Y2 = Y[,2]
  
  eta = min(Y) - 1 - time - 2 *(max(Y)-min(Y)) # in order to be worst
  distmin = min(abs(Y1-Y2))/time/10 # less than minimum difference between Y1 and Y2 divided by time
  
  Y1_tilde = Y1 * (t1 >= time) + (eta + t1*distmin) * (t1 < time)
  Y2_tilde = Y2 * (t2 >= time) + (eta + t2*distmin) * (t2 < time)
  
  Y1_survivorOnly = Y1[t1 >= time & t2 >= time]
  Y2_survivorOnly = Y2[t1 >= time & t2 >= time]
  
  OnlyControlDead = sum(t1 < time & t2 >= time)
  OnlyTreatedDead = sum(t2 < time & t1 >= time)
  
  return(list(Y1_tilde, Y2_tilde, Y1_survivorOnly, Y2_survivorOnly, OnlyControlDead, OnlyTreatedDead))
}  

calculate.binom.pval <- function(OnlyControlDead, OnlyTreatedDead, beta) {
  # T1 : binomial
  pval = binom.test(c(OnlyControlDead, OnlyTreatedDead), alternative = "greater")$p.value;
  if (pval <= beta) return (pval)
  else { 
    if (OnlyTreatedDead == 0) {pval.alt = 0}
    else {pval.alt = binom.test(c(OnlyControlDead+1, OnlyTreatedDead-1), alternative = "greater")$p.value}
    
    if (pval.alt > beta) {return (pval)}
    # auxiliary random 
    else {return (1 - rbinom(1, 1, (beta - pval.alt)/(pval - pval.alt)))}
  }
}
```


```{r}
power.exp.seqT1T2 = array(data = NA, dim=c(length(q2), length(Delta), length(HR), length(beta)), dimnames = list(c("q2=60%", "q2=80%", "q2=90%"), c(0.0, 0.1, 0.2, 0.3), c(1.0, 1.2, 1.4, 1.6), seq(0, 0.05, by=0.001)))

power.exp.seqT1T3 = array(data = NA, dim=c(length(q2), length(Delta), length(HR), length(beta)), dimnames = list(c("q2=60%", "q2=80%", "q2=90%"), c(0.0, 0.1, 0.2, 0.3), c(1.0, 1.2, 1.4, 1.6), seq(0, 0.05, by=0.001)))

for (i in 1:length(q2)) {
  cat("### q2:", q2[i], "\n")
  for (j in 1:length(Delta)) {
    cat("## Delta:", Delta[j], "\n")
    for (k in 1:length(HR)) {
      cat("# HR:", HR[k], "\n")
      count12 = array(0, length(beta))
      count13 = array(0, length(beta))
      total = 0
      for (l in c(1:N)) {
        outcomes = generate.outcomes.exp(n, q2[i], HR[k], Delta[j], time, rho_Y, rho_t)
        Y1_tilde = outcomes[[1]]
        Y2_tilde = outcomes[[2]]
        Y1_survivorOnly = outcomes[[3]]
        Y2_survivorOnly = outcomes[[4]]
        OnlyControlDead = outcomes[[5]]
        OnlyTreatedDead = outcomes[[6]]
        # if every pair has dead person, pass
        if (length(Y1_survivorOnly) == 0) next;
        # if there's no discordant pair, pass
        if (OnlyControlDead == 0 & OnlyTreatedDead == 0) next;
        total = total + 1 # denominator
        
        # T2 = worst rank
        pval.wr = wilcox.test(Y1_tilde, Y2_tilde, alternative="less", paired=T)$p.value
        
        # T3 = survivors only
        pval.surv = wilcox.test(Y1_survivorOnly, Y2_survivorOnly, alternative = "less", paired = T)$p.value 
        
        for (b in 1:length(beta)) {
          pval.binom = calculate.binom.pval(OnlyControlDead, OnlyTreatedDead, b)

          if (pval.binom <= beta[b]) {
            count12[b] = count12[b] + 1
            count13[b] = count13[b] + 1
          }
          else {
            if (pval.wr <= (alpha - beta[b])) {count12[b] = count12[b] + 1}
            if (pval.surv <= (alpha - beta[b])) {count13[b] = count13[b] + 1}
          }
        }
      }
      cat("total missing =", N - total, "\n")
      power.exp.seqT1T2[i,j,k,] = count12/total
      power.exp.seqT1T3[i,j,k,] = count13/total
    }
  }
}
```

#### plot

```{r}
exp.seqT1T2.q2low = power.exp.seqT1T2[1,,,]
exp.seqT1T3.q2low = power.exp.seqT1T3[1,,,]

exp.seqT1T2.q2mid = power.exp.seqT1T2[2,,,]
exp.seqT1T3.q2mid = power.exp.seqT1T3[2,,,]

exp.seqT1T2.q2high = power.exp.seqT1T2[3,,,]
exp.seqT1T3.q2high = power.exp.seqT1T3[3,,,]
```


##### (1) q2 = low (60%)

```{r}
power12_delta0_hr1.0 = exp.seqT1T2.q2low[1,1,] # Delta=0, HR=1
power12_delta0_hr1.2 = exp.seqT1T2.q2low[1,2,] # Delta=0, HR=1.2
power12_delta0_hr1.4 = exp.seqT1T2.q2low[1,3,] # Delta=0, HR=1.4
power12_delta0_hr1.6 = exp.seqT1T2.q2low[1,4,] # Delta=0, HR=1.6

power12_delta0.1_hr1.0 = exp.seqT1T2.q2low[2,1,] # Delta=0.1, HR=1
power12_delta0.1_hr1.2 = exp.seqT1T2.q2low[2,2,] # Delta=0.1, HR=1.2
power12_delta0.1_hr1.4 = exp.seqT1T2.q2low[2,3,] # Delta=0.1, HR=1.4
power12_delta0.1_hr1.6 = exp.seqT1T2.q2low[2,4,] # Delta=0.1, HR=1.6

power12_delta0.2_hr1.0 = exp.seqT1T2.q2low[3,1,] # Delta=0.2, HR=1
power12_delta0.2_hr1.2 = exp.seqT1T2.q2low[3,2,] # Delta=0.2, HR=1.2
power12_delta0.2_hr1.4 = exp.seqT1T2.q2low[3,3,] # Delta=0.2, HR=1.4
power12_delta0.2_hr1.6 = exp.seqT1T2.q2low[3,4,] # Delta=0.2, HR=1.6

power12_delta0.3_hr1.0 = exp.seqT1T2.q2low[4,1,] # Delta=0.3, HR=1
power12_delta0.3_hr1.2 = exp.seqT1T2.q2low[4,2,] # Delta=0.3, HR=1.2
power12_delta0.3_hr1.4 = exp.seqT1T2.q2low[4,3,] # Delta=0.3, HR=1.4
power12_delta0.3_hr1.6 = exp.seqT1T2.q2low[4,4,] # Delta=0.3, HR=1.6

power13_delta0_hr1.0 = exp.seqT1T3.q2low[1,1,] # Delta=0, HR=1
power13_delta0_hr1.2 = exp.seqT1T3.q2low[1,2,] # Delta=0, HR=1.2
power13_delta0_hr1.4 = exp.seqT1T3.q2low[1,3,] # Delta=0, HR=1.4
power13_delta0_hr1.6 = exp.seqT1T3.q2low[1,4,] # Delta=0, HR=1.6

power13_delta0.1_hr1.0 = exp.seqT1T3.q2low[2,1,] # Delta=0.1, HR=1
power13_delta0.1_hr1.2 = exp.seqT1T3.q2low[2,2,] # Delta=0.1, HR=1.2
power13_delta0.1_hr1.4 = exp.seqT1T3.q2low[2,3,] # Delta=0.1, HR=1.4
power13_delta0.1_hr1.6 = exp.seqT1T3.q2low[2,4,] # Delta=0.1, HR=1.6

power13_delta0.2_hr1.0 = exp.seqT1T3.q2low[3,1,] # Delta=0.2, HR=1
power13_delta0.2_hr1.2 = exp.seqT1T3.q2low[3,2,] # Delta=0.2, HR=1.2
power13_delta0.2_hr1.4 = exp.seqT1T3.q2low[3,3,] # Delta=0.2, HR=1.4
power13_delta0.2_hr1.6 = exp.seqT1T3.q2low[3,4,] # Delta=0.2, HR=1.6

power13_delta0.3_hr1.0 = exp.seqT1T3.q2low[4,1,] # Delta=0.3, HR=1
power13_delta0.3_hr1.2 = exp.seqT1T3.q2low[4,2,] # Delta=0.3, HR=1.2
power13_delta0.3_hr1.4 = exp.seqT1T3.q2low[4,3,] # Delta=0.3, HR=1.4
power13_delta0.3_hr1.6 = exp.seqT1T3.q2low[4,4,] # Delta=0.3, HR=1.6

df = data.frame(beta = rep(beta, each=16),
                HR = as.factor(rep(rep(c(1.0, 1.2, 1.4, 1.6), each=4), length(beta))),
                delta = as.factor(rep(c(0, 0.1, 0.2, 0.3), 4*length(beta))),
                power12 = c(exp.seqT1T2.q2low),
                power13 = c(exp.seqT1T3.q2low))
#power12 = 
#c(power12_delta0_hr1.0, power12_delta0.025_hr1.0, power12_delta0.05_hr1.0, power12_delta0.1_hr1.0,
#  power12_delta0_hr1.1, power12_delta0.025_hr1.1, power12_delta0.05_hr1.1, power12_delta0.1_hr1.1,
#  power12_delta0_hr1.2, power12_delta0.025_hr1.2, power12_delta0.05_hr1.2, power12_delta0.1_hr1.2,
#  power12_delta0_hr1.3, power12_delta0.025_hr1.3, power12_delta0.05_hr1.3, power12_delta0.1_hr1.3), 
#power13 = 
#c(power13_delta0_hr1.0, power13_delta0.025_hr1.0, power13_delta0.05_hr1.0, power13_delta0.1_hr1.0,
#  power13_delta0_hr1.1, power13_delta0.025_hr1.1, power13_delta0.05_hr1.1, power13_delta0.1_hr1.1,
#  power13_delta0_hr1.2, power13_delta0.025_hr1.2, power13_delta0.05_hr1.2, power13_delta0.1_hr1.2,
#  power13_delta0_hr1.3, power13_delta0.025_hr1.3, power13_delta0.05_hr1.3, power13_delta0.1_hr1.3))

#print(df)
```

- find $argmax_{\beta}$ for each HR and delta

```{r}
for (hr in unique(df$HR)) {
  for (d in unique(df$delta)) {
    gp = df %>% filter(HR==hr) %>% filter(delta==d)
    max_beta12 = gp$beta[which(gp$power12 == max(gp$power12))]
    max_beta13 = gp$beta[which(gp$power13 == max(gp$power13))]
    cat("argmax beta for T1+T2 / HR", hr, "/ delta", d, "=", max_beta12, "\n")
    cat("argmax beta for T1+T3 / HR", hr, "/ delta", d, "=", max_beta13, "\n")
  }
}
```


```{r}
g1.0low <- df %>% filter(HR==1.0) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 60%, HR = 1.0") 

g1.2low <- df %>% filter(HR==1.2) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 60%, HR = 1.2") 

g1.4low <- df %>% filter(HR==1.4) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 60%, HR = 1.4") 

g1.6low <- df %>% filter(HR==1.6) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 60%, HR = 1.6") 

g1.0low
g1.2low
g1.4low
g1.6low
```


##### (2) q2 = mid (80%)

```{r}
power12_delta0_hr1.0 = exp.seqT1T2.q2mid[1,1,] # Delta=0, HR=1
power12_delta0_hr1.2 = exp.seqT1T2.q2mid[1,2,] # Delta=0, HR=1.2
power12_delta0_hr1.4 = exp.seqT1T2.q2mid[1,3,] # Delta=0, HR=1.4
power12_delta0_hr1.6 = exp.seqT1T2.q2mid[1,4,] # Delta=0, HR=1.6

power12_delta0.1_hr1.0 = exp.seqT1T2.q2mid[2,1,] # Delta=0.1, HR=1
power12_delta0.1_hr1.2 = exp.seqT1T2.q2mid[2,2,] # Delta=0.1, HR=1.2
power12_delta0.1_hr1.4 = exp.seqT1T2.q2mid[2,3,] # Delta=0.1, HR=1.4
power12_delta0.1_hr1.6 = exp.seqT1T2.q2mid[2,4,] # Delta=0.1, HR=1.6

power12_delta0.2_hr1.0 = exp.seqT1T2.q2mid[3,1,] # Delta=0.2, HR=1
power12_delta0.2_hr1.2 = exp.seqT1T2.q2mid[3,2,] # Delta=0.2, HR=1.2
power12_delta0.2_hr1.4 = exp.seqT1T2.q2mid[3,3,] # Delta=0.2, HR=1.4
power12_delta0.2_hr1.6 = exp.seqT1T2.q2mid[3,4,] # Delta=0.2, HR=1.6

power12_delta0.3_hr1.0 = exp.seqT1T2.q2mid[4,1,] # Delta=0.3, HR=1
power12_delta0.3_hr1.2 = exp.seqT1T2.q2mid[4,2,] # Delta=0.3, HR=1.2
power12_delta0.3_hr1.4 = exp.seqT1T2.q2mid[4,3,] # Delta=0.3, HR=1.4
power12_delta0.3_hr1.6 = exp.seqT1T2.q2mid[4,4,] # Delta=0.3, HR=1.6

power13_delta0_hr1.0 = exp.seqT1T3.q2mid[1,1,] # Delta=0, HR=1
power13_delta0_hr1.2 = exp.seqT1T3.q2mid[1,2,] # Delta=0, HR=1.2
power13_delta0_hr1.4 = exp.seqT1T3.q2mid[1,3,] # Delta=0, HR=1.4
power13_delta0_hr1.6 = exp.seqT1T3.q2mid[1,4,] # Delta=0, HR=1.6

power13_delta0.1_hr1.0 = exp.seqT1T3.q2mid[2,1,] # Delta=0.1, HR=1
power13_delta0.1_hr1.2 = exp.seqT1T3.q2mid[2,2,] # Delta=0.1, HR=1.2
power13_delta0.1_hr1.4 = exp.seqT1T3.q2mid[2,3,] # Delta=0.1, HR=1.4
power13_delta0.1_hr1.6 = exp.seqT1T3.q2mid[2,4,] # Delta=0.1, HR=1.6

power13_delta0.2_hr1.0 = exp.seqT1T3.q2mid[3,1,] # Delta=0.2, HR=1
power13_delta0.2_hr1.2 = exp.seqT1T3.q2mid[3,2,] # Delta=0.2, HR=1.2
power13_delta0.2_hr1.4 = exp.seqT1T3.q2mid[3,3,] # Delta=0.2, HR=1.4
power13_delta0.2_hr1.6 = exp.seqT1T3.q2mid[3,4,] # Delta=0.2, HR=1.6

power13_delta0.3_hr1.0 = exp.seqT1T3.q2mid[4,1,] # Delta=0.3, HR=1
power13_delta0.3_hr1.2 = exp.seqT1T3.q2mid[4,2,] # Delta=0.3, HR=1.2
power13_delta0.3_hr1.4 = exp.seqT1T3.q2mid[4,3,] # Delta=0.3, HR=1.4
power13_delta0.3_hr1.6 = exp.seqT1T3.q2mid[4,4,] # Delta=0.3, HR=1.6

df = data.frame(beta = rep(beta, each=16),
                HR = as.factor(rep(rep(c(1.0, 1.2, 1.4, 1.6), each=4), length(beta))),
                delta = as.factor(rep(c(0, 0.1, 0.2, 0.3), 4*length(beta))),
                power12 = c(exp.seqT1T2.q2mid),
                power13 = c(exp.seqT1T3.q2mid))
```

- find $argmax_{\beta}$ for each HR and delta

```{r}
for (hr in unique(df$HR)) {
  for (d in unique(df$delta)) {
    gp = df %>% filter(HR==hr) %>% filter(delta==d)
    max_beta12 = gp$beta[which(gp$power12 == max(gp$power12))]
    max_beta13 = gp$beta[which(gp$power13 == max(gp$power13))]
    cat("argmax beta for T1+T2 / HR", hr, "/ delta", d, "=", max_beta12, "\n")
    cat("argmax beta for T1+T3 / HR", hr, "/ delta", d, "=", max_beta13, "\n")
  }
}
```


```{r}
g1.0mid <- df %>% filter(HR==1.0) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 80%, HR = 1.0") 

g1.2mid <- df %>% filter(HR==1.2) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 80%, HR = 1.2") 

g1.4mid <- df %>% filter(HR==1.4) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 80%, HR = 1.4") 

g1.6mid <- df %>% filter(HR==1.6) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 80%, HR = 1.6") 

g1.0mid
g1.2mid
g1.4mid
g1.6mid
```


##### (3) q2 = high (90%)

```{r}
power12_delta0_hr1.0 = exp.seqT1T2.q2high[1,1,] # Delta=0, HR=1
power12_delta0_hr1.2 = exp.seqT1T2.q2high[1,2,] # Delta=0, HR=1.2
power12_delta0_hr1.4 = exp.seqT1T2.q2high[1,3,] # Delta=0, HR=1.4
power12_delta0_hr1.6 = exp.seqT1T2.q2high[1,4,] # Delta=0, HR=1.6

power12_delta0.1_hr1.0 = exp.seqT1T2.q2high[2,1,] # Delta=0.1, HR=1
power12_delta0.1_hr1.2 = exp.seqT1T2.q2high[2,2,] # Delta=0.1, HR=1.2
power12_delta0.1_hr1.4 = exp.seqT1T2.q2high[2,3,] # Delta=0.1, HR=1.4
power12_delta0.1_hr1.6 = exp.seqT1T2.q2high[2,4,] # Delta=0.1, HR=1.6

power12_delta0.2_hr1.0 = exp.seqT1T2.q2high[3,1,] # Delta=0.2, HR=1
power12_delta0.2_hr1.2 = exp.seqT1T2.q2high[3,2,] # Delta=0.2, HR=1.2
power12_delta0.2_hr1.4 = exp.seqT1T2.q2high[3,3,] # Delta=0.2, HR=1.4
power12_delta0.2_hr1.6 = exp.seqT1T2.q2high[3,4,] # Delta=0.2, HR=1.6

power12_delta0.3_hr1.0 = exp.seqT1T2.q2high[4,1,] # Delta=0.3, HR=1
power12_delta0.3_hr1.2 = exp.seqT1T2.q2high[4,2,] # Delta=0.3, HR=1.2
power12_delta0.3_hr1.4 = exp.seqT1T2.q2high[4,3,] # Delta=0.3, HR=1.4
power12_delta0.3_hr1.6 = exp.seqT1T2.q2high[4,4,] # Delta=0.3, HR=1.6

power13_delta0_hr1.0 = exp.seqT1T3.q2high[1,1,] # Delta=0, HR=1
power13_delta0_hr1.2 = exp.seqT1T3.q2high[1,2,] # Delta=0, HR=1.2
power13_delta0_hr1.4 = exp.seqT1T3.q2high[1,3,] # Delta=0, HR=1.4
power13_delta0_hr1.6 = exp.seqT1T3.q2high[1,4,] # Delta=0, HR=1.6

power13_delta0.1_hr1.0 = exp.seqT1T3.q2high[2,1,] # Delta=0.1, HR=1
power13_delta0.1_hr1.2 = exp.seqT1T3.q2high[2,2,] # Delta=0.1, HR=1.2
power13_delta0.1_hr1.4 = exp.seqT1T3.q2high[2,3,] # Delta=0.1, HR=1.4
power13_delta0.1_hr1.6 = exp.seqT1T3.q2high[2,4,] # Delta=0.1, HR=1.6

power13_delta0.2_hr1.0 = exp.seqT1T3.q2high[3,1,] # Delta=0.2, HR=1
power13_delta0.2_hr1.2 = exp.seqT1T3.q2high[3,2,] # Delta=0.2, HR=1.2
power13_delta0.2_hr1.4 = exp.seqT1T3.q2high[3,3,] # Delta=0.2, HR=1.4
power13_delta0.2_hr1.6 = exp.seqT1T3.q2high[3,4,] # Delta=0.2, HR=1.6

power13_delta0.3_hr1.0 = exp.seqT1T3.q2high[4,1,] # Delta=0.3, HR=1
power13_delta0.3_hr1.2 = exp.seqT1T3.q2high[4,2,] # Delta=0.3, HR=1.2
power13_delta0.3_hr1.4 = exp.seqT1T3.q2high[4,3,] # Delta=0.3, HR=1.4
power13_delta0.3_hr1.6 = exp.seqT1T3.q2high[4,4,] # Delta=0.3, HR=1.6

df = data.frame(beta = rep(beta, each=16),
                HR = as.factor(rep(rep(c(1.0, 1.2, 1.4, 1.6), each=4), length(beta))),
                delta = as.factor(rep(c(0, 0.1, 0.2, 0.3), 4*length(beta))),
                power12 = c(exp.seqT1T2.q2high),
                power13 = c(exp.seqT1T3.q2high))
```

- find $argmax_{\beta}$ for each HR and delta

```{r}
for (hr in unique(df$HR)) {
  for (d in unique(df$delta)) {
    gp = df %>% filter(HR==hr) %>% filter(delta==d)
    max_beta12 = gp$beta[which(gp$power12 == max(gp$power12))]
    max_beta13 = gp$beta[which(gp$power13 == max(gp$power13))]
    cat("argmax beta for T1+T2 / HR", hr, "/ delta", d, "=", max_beta12, "\n")
    cat("argmax beta for T1+T3 / HR", hr, "/ delta", d, "=", max_beta13, "\n")
  }
}
```


```{r}
g1.0high <- df %>% filter(HR==1.0) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 90%, HR = 1.0") 

g1.2high <- df %>% filter(HR==1.2) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 90%, HR = 1.2") 

g1.4high <- df %>% filter(HR==1.4) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 90%, HR = 1.4") 

g1.6high <- df %>% filter(HR==1.6) %>%
  ggplot() +
  geom_line(mapping = aes(x=beta, y=power12, group=delta, col=delta), linetype=1, size=1) +
  geom_line(mapping = aes(x=beta, y=power13, group=delta, col=delta), linetype=6, size=1) +
  theme_bw() +
  scale_y_continuous(limits=c(0, 1)) +
  ylab("power") +
  ggtitle("q2 = 90%, HR = 1.6") 

g1.0high
g1.2high
g1.4high
g1.6high
```




### Plot at once

```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
shared_legend <- extract_legend(g1.0high)

g1.0low <- g1.0low + theme(legend.position="none")
g1.2low <- g1.2low + theme(legend.position="none")
g1.4low <- g1.4low + theme(legend.position="none")
g1.6low <- g1.6low + theme(legend.position="none")
g1.0mid <- g1.0mid + theme(legend.position="none")
g1.2mid <- g1.2mid + theme(legend.position="none")
g1.4mid <- g1.4mid + theme(legend.position="none")
g1.6mid <- g1.6mid + theme(legend.position="none")
g1.0high <- g1.0high + theme(legend.position="none")
g1.2high <- g1.2high + theme(legend.position="none")
g1.4high <- g1.4high + theme(legend.position="none")
g1.6high <- g1.6high + theme(legend.position="none")

grid.arrange(g1.0low, g1.2low, g1.4low, g1.6low, 
             g1.0mid, g1.2mid, g1.4mid, g1.6mid,
             g1.0high, g1.2high, g1.4high, g1.6high,
             shared_legend, ncol=5, widths=c(3,3,3,3,1),
             layout_matrix=cbind(c(1,5,9),c(2,6,10),c(3,7,11),c(4,8,12),c(13,13,13)))
```